Clustering feature points

### 2013-05-29 text/clustering.tex
Added stuff to clustering and new results

\section{Project description}

Traditionally the feature points of two images have been matched by 
comparing the feature points individually and finding the best matches 
based on their distance. This approach was originally introduced with 
the SIFT features \cite{lowe2004sift} and has been used since then in 
computer vision applications such as scene recognition\footnote{See for 
example chapter 4.1.3 of \cite[pp. 226]{szeliski2010}} and image 
alignment\footnote{An example is the gold standard algorithm in \cite[p.  
114]{multipleView}}. In larger scale approaches where an image is 
compared with thousands or millions of other images, the feature points 
are usually indexed in a dictionary of visual words instead. This 
approach makes it possible to quickly retrieve matches in a large set of 
potential candidates, but the performance is restricted by how 
discriminative the individual feature points are. To agument the 
descriminative power of the feature points, \cite{wu2009bundling} 
proposes grouping the features and use the geometrical relationships 
between members of a group when matching two feature points. Similarly 
\cite{das2008event} use geometrical information about the feature points 
to cluster them before doing scene matching resulting in better matches.

@@

Since the introduction of the SIFT\cite features{lowe2004sift}, 
collecting feature points from an image and using the descriptors to 
match images have been used in a plethora of different applications. The 
development of feature points have not been standing still since Lowe 
introduced the SIFT feature back in 2004. Various modifications and 
improments both to how features are collected and how features are best 
described have been suggested since then. In locating feature points in 
an image the focus

>>>>

\section{Introduction}

Since the introduction of feature points, several methods for matching 
them across images have been proposed in order to maximize the amount of 
correct matches. In this part of the project I introduce two methods to 
reliably match feature points. The word 'reliable' covers that matches 
only are returned if they are judged as sound. That is, matches should 
only be returned if there are objects or scenes that appear in both 
images.

When matching feature points between two images we are often faced with 
a dilemma between making the feature descriptor more discriminative to 
make sure only correct matches are made, and making it less 
discriminative to increase the robustness of the descriptor over 
different changes such as illumination, perspective and image quality.  
Any feature type aiming for robustness will have cases where it isn't 
possible to generate reliable matches if we only look at the pairwise 
correspondence of the descriptors.

To get around this limitation several methods have been proposed in the 
literature. One of the most fruitful approaches have been to look at the 
geometric configuration of the feature points and match them with 
respect to assumptions concerning the variance between images. For 
example an angular and distance constraint between matches as considered 
by \cite{kim2008efficient} performs well in situations where no camera 
rotation occurs between the two images but will likely fail in other 
cases. Various scenarios have been proposed to improve on this simple 
assumption, such as epipolar constraints (\cite{torr2000mlesac}, 
\cite{chum2005matching}) and pairwise constraints 
(\cite{choi2009robust}, \cite{leordenanu2005spectral}). The Epipolar 
constraint carries the assumption that the two images matched are 
related by an affine transformation. That is, there is no relative 
movement of objects in between images and either the viewpoint is fixed 
or the image resides entirely on a plane. In practice this assumption 
largely holds true when all objects we are interested in matching are 
roughly the same distance from the camera and when we don't expect to 
match objects that aren't consistently positioned across images.  
Pairwise constraints provide a more robust approach to this problem by 
looking at a set of proposed correspondences and defining a pairwise 
error between any two matches usually based on the assumption that two 
neighboring correspondences will usually have similar angles and 
distances. We can then convert the problem to an optimization problem 
and return a set of correspondences that minimizes this error such as 
proposed in \cite{choi2009robust} and \cite{leordeanu2005spectral}.  
This approach provides more robustness because assumptions that might 
not be true globally often still holds on a local level.

A similar approach is to pick out different zones in each image and

In practice there are many situations where using the geometry of the 
image to filter correspondences is not possible. This situation most 
often arise when the assumptions made by a geometric method turns out 
not to hold in a particular use case. This could matchings between 
images where the content has been scrambled or with so little 
consistency that even the geometric assumptions break down even on a 
local level. In addition the geometric methods all require a set of 
correspondences to begin with. If this set of correspondences is 
narrowed down to a smaller set with a higher ratio of inliers the result 
of the geometric matching will be both faster and more accurate. Finally 
some use cases might require a performance that can't be achieved by a 
more complex geometric method where simple non geometric methods might 
be able to provide additional matching accuracy with a smaller 
performance penalty.

There are a relatively small number of algorithms proposed to solve the 
matching problem without involving geometric constraints. Traditionally 
the feature points of two images have been matched by comparing the 
every feature point of one image with all feature points of the other 
and finding the best matches based on the similarity of the descriptors.  
With the introduction of the SIFT features \cite{lowe2004sift}, Lowe 
proposes an alternative measure where the uniqueness of a given match is 
found by looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities. By ranking the 
scores by their uniqueness and picking the $n$ best we get a set of 
correspondences that are distinctly matched across the two images. In 
the book Multiple View Geometry \cite[p.  114]{multipleView}, the gold 
standard algorithm is introduced which ranks correspondences by 
similarity and further filters them by making sure that the map of 
matches is reflective, i.e. that for every feature point $f_i$ matching 
$f_j$ we also have that $f_j$ matches $f_i$.

In turn the two methods proposed in this report are inspired by a simple 
but novel idea. If we have two images and a given feature point in the 
first image is better matched with other feature points from the 
\emph{same} image than points in the other image, then any matches of 
this feature point to points in the other image is deemed unreliable and 
discarded.  This approach carries no implicit assumptions about the 
geometric consistency of matches and can as such easily be extended with 
other geometric solutions when appropriate.

Based on this idea the two proposed methods find reliable matches as 
follows:
\begin{itemize}
\item{\emph{Mirror Match}: Take any traditional way of matching feature 
but extend it to match any given feature with the combined set of 
features from both images, filtering out all matches that are within the 
same image.}
\item{\emph{Cluster Match}: Take the combined set of feature points from 
both images and cluster these points according to their
descriptors. Given a resulting partition of points, no matches are 
returned if it contains only feature points from one image. If the 
partition contains points from both images, \emph{Mirror Match} is used 
to find the best matches within the partition.
\end{itemize}
The

%Given two sets of feature points, $A$ and $B$ corresponding to two 
%images $i_A$ and $i_B$, and a feature point $f_A^i$ in set $A$ is more 
%closely related to one or more other feature points in set $A$ than 
%feature points in set $B$ then the potential matches between $f_A^i$ 
%and feature points in $B$ is deemed \emph{unreliable}.

 This approach was originally introduced with the SIFT features and has 
been used since then in computer vision applications such as scene 
recognition\footnote{See for example chapter 4.1.3 of \cite[pp.  
226]{szeliski2010}} and image alignment\footnote{An example is the gold 
standard algorithm in \cite[p.  114]{multipleView}}.  In larger scale 
approaches where an image is compared with thousands or millions of 
other images, the feature points are usually indexed in a dictionary of 
visual words instead. This approach makes it possible to quickly 
retrieve matches in a large set of potential candidates, but the 
performance is restricted by how discriminative the individual feature 
points are. To agument the descriminative power of the feature points, 
\cite{wu2009bundling} proposes grouping the features and use the 
geometrical relationships between members of a group when matching two 
feature points. Similarly \cite{das2008event} use geometrical 
information about the feature points to cluster them before doing scene 
matching resulting in better matches.

@@

Since the introduction of the SIFT features\cite{lowe2004sift}, 
collecting feature points from an image and using the descriptors to 
match images have been used in a plethora of different applications. The 
development of feature points have not been standing still since Lowe 
introduced the SIFT feature back in 2004. Various modifications and 
improments both to how features are collected and how features are best 
described have been suggested since then. In locating feature points in 
an image the focus
